---
title: "r2dii.plot"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of r2dii.plot is to help you plot 2DII data in an informative,
beautiful, and easy way. It is designed to work smoothly with other "r2dii"
packages -- [r2dii.data](https://2degreesinvesting.github.io/r2dii.data/),
[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/), and
[r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/). It also
plays well with the [ggplot2](https://ggplot2.tidyverse.org/) package, which
helps you customize your plots.

```{r setup}
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
library(r2dii.data)
library(r2dii.match)
library(r2dii.analysis)
library(r2dii.plot)
```

Your data should have a structure similar to that of the demo datasets in the
r2dii.data package.

```{r}
loanbook <- loanbook_demo
ald <- ald_demo
scenario <- co2_intensity_scenario_demo
region <- region_isos_demo
```

Your workflow involves functions from the packages r2dii.match and
r2dii.analysis.

```{r}
matched <- loanbook %>%
  match_name(ald) %>%
  prioritize() # Remember to validate matches (see `?prioritize`)
```

r2dii.plot supports three kinds of plots:

* Emission intensity,
* Techmix,
* Volume trajectory.

Each plot can be created using two types of functions: 

* `qplot_*()` for standardized plots,
* `plot_*()` for 'bare' plots leaving the customization
to the user.

For each kind, you'll need to subset the specific rows you want to plot (for
example with `subset()` or `dplyr:filter()`). For details see the documented
"Requirements" of the argument `data` of each `plot_*()` or `qplot_*()` function
(e.g. see `?plot_emission_intensity`). If you forget to meet the `data`
requirements the error message should guide you.

## 'Quick' plots

Quick plots allow you to create a good looking plot without extensive knowledge
of other R packages. They are perfect to have a quick look into the data and for
users who are interested in standard PACTA visualizations. Their common
characteristics are:

* 5 year time span.
* Formatted data labels.
* Formatted axis labels.
* A title (and a subtitle).

Use `qplot_emission_intensity()` with `sda`-like data.

```{r}
data <- matched %>%
  target_sda(ald, co2_intensity_scenario = scenario, region_isos = region) %>%
  filter(sector == "cement")

qplot_emission_intensity(data)
```

Use `qplot_trajectory()` with `market_share`-like data.

```{r}
data <- matched %>%
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) %>%
  filter(technology == "renewablescap", region == "global")

qplot_trajectory(data)
```

Use `qplot_techmix()` with `market_share`-like data.

```{r}
data <- matched %>%
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) %>%
  filter(
    sector == "power",
    region == "global",
    metric %in% c("projected", "corporate_economy", "target_sds")
  )

qplot_techmix(data)
```

## Plots

Plots created with `plot_*()` functions show the data as they are. To customize
these plots you can use three strategies: 

1) Use parameters of `plot_*()` functions, 
2) Modify the input `data`,
3) Use `ggplot2` functions.

The following sub sections show the three strategies applied to different functions 
of the `plot_*()` family.

### Plot customization strategies 1 and 3: parameters and ggplot2 functions

We will show how you might customize a plot using strategy 1 and 3 based on 
`plot_emission_intensity()`. 

The basic output of the function looks rather unappealing.

```{r}
data <- matched %>%
  target_sda(ald, co2_intensity_scenario = scenario, region_isos = region) %>%
  filter(sector == "cement")

plot_emission_intensity(data)
```

You can use parameters of `plot_emission_intensity()` to replicate features of 
`qplot_emission_intensity()`(strategy 1), and add plot labels with 
`ggplot2::labs()` (strategy 3).

```{r}
data <- matched %>%
  target_sda(ald, co2_intensity_scenario = scenario, region_isos = region) %>%
  filter(sector == "cement")

plot_emission_intensity(data, convert_label = to_title, span_5yr = TRUE) +
  labs(
    title = "Emission Intensity Trend for the Cement Sector",
    x = "Year",
    y = "Tons of CO2 per Ton of Production Unit"
  )
```

### Plot customization strategies 2 and 3: input data and ggplot2 functions

You can also polish your plot by modifying the input `data` (strategy 2), for 
example: 

* Changing the time span.
* Adding custom labels by adding a column 'label' or 'label_tech' to
`market_share`-like data.

And by modifying output `ggplot` object (strategy 3), for example:

* Adding a title and a subtitle using `ggplot2::labs()`. 
* Changing x and y axis labels using `ggplot2::labs()`.
* Customizing the colours and legend labels with `ggplot2::scale_colour_manual()`
or `r2dii.plot::scale_*()` functions (see [Styling functions]).

Here is how you might customize each of the three kinds of plots using strategy 
2 and 3:

* `plot_emission_intensity()` - add custom title and axis labels and modify 
colours and legend labels.

```{r}
data <- sda %>%
  filter(
    sector == "cement",
    year <= 2030
  )

plot_emission_intensity(data) +
  labs(
    title = "Emission intensity plot for cement",
    x = "Time",
    y = "Tons of CO2 per ton of cement produced"
  ) +
  scale_color_manual(
    values = c("#4a5e54", "#a63d57", "#78c4d6", "#f2e06e"),
    labels = c("Proj.", "Corp. Economy", "Target (demo)", "Adj. Scenario (demo)")
  )
```

* `plot_trajectory()` - change time span, add 'label' column and add custom 
title and axis labels.

```{r}
data <- matched %>%
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) %>%
  filter(
    technology == "renewablescap",
    region == "global",
    year <= 2030
  ) %>%
  mutate(
    label = case_when(
      metric == "projected" ~ "Your Portfolio",
      metric == "corporate_economy" ~ "Benchmark (Corp. Economy)",
      metric == "target_sds" ~ "Sustainable Development Scen.",
      metric == "target_sps" ~ "Stated Policies Scen.",
      metric == "target_cps" ~ "Current Policy Scen.",
      TRUE ~ metric
    )
  )

plot_trajectory(data) +
  scale_x_continuous(n.breaks = 3) +
  labs(
    title = "Portfolio Scenario Alignment for Renewables Technology",
    x = "Year",
    y = "Production normalized to start year"
  ) +
  theme(plot.margin = unit(c(0.5, 6, 0.5, 1), "cm")) #so the long labels fit
```

* `plot_techmix()` - change time span, add 'label' column, apply custom colours 
and modify legend labels.

```{r}
data <- market_share %>%
  filter(
    metric %in% c("projected", "corporate_economy", "target_sds"),
    sector == "power",
    region == "global",
    year >= 2021,
    year <= 2040 # custom time range
  ) %>%
  mutate(
    label = case_when(
      metric == "projected" ~ "Your Portfolio",
      metric == "corporate_economy" ~ "Corporate Economy Benchmark",
      metric == "target_sds" ~ "SDS Scenario"
    )
  )

plot_techmix(data) +
  scale_fill_manual(
    values = c("black", "brown", "grey", "yellow", "blue", "green4"),
    labels = paste(c("Coal", "Oil", "Gas", "Nuclear", "Hydro", "Renewables"), "Cap.")
  )
```

* `plot_techmix()` - remove scenario data for start year.

```{r}
data <- market_share %>%
  filter(
    metric %in% c("projected", "corporate_economy", "target_sds"),
    sector == "power",
    region == "global"
  ) %>%
  filter(
    !((metric == "target_sds") & (year == 2020))
  )

plot_techmix(data)
```

## Styling functions

A number of functions allow you to customize any of your plots using the 2DII style:

* Use `theme_2dii()` to change the display of non-data content.

```{r}
data <- market_share %>%
  filter(
    metric == "projected",
    sector == "power",
    region == "global",
    year %in% c(2020, 2025)
  )

ggplot(data, aes(x = factor(year), y = production)) +
  geom_col() +
  facet_wrap(~technology) +
  theme_2dii()
```

* Use `scale_colour_r2dii()` and `scale_fill_r2dii()` to apply the 2DII colour
palette (see `?scale_colour_2dii()` to find out what are the available colour 
labels).

```{r}
data <- market_share %>%
  filter(
    metric != "corporate_economy",
    sector == "power",
    region == "global",
    technology == "renewablescap"
  )

ggplot(data, aes(x = year, y = production, color = metric)) +
  geom_line() +
  scale_colour_r2dii(labels = c("dark_blue", "green", "orange", "ruby_red")) +
  theme_2dii()
```

* Use `scale_colour_r2dii_sector()` and `scale_fill_r2dii_sector()` to apply
the 2DII colour palette for sectors (see `?scale_colour_2dii_sector()` to find 
out what are the available colour labels).

```{r}
data <- market_share %>%
  filter(
    metric == "projected",
    region == "global",
    year %in% c(2020, 2025)
  ) %>%
  group_by(sector, year) %>%
  summarise(production = sum(production))

ggplot(data, aes(x = factor(year), y = production, fill = sector)) +
  geom_col() +
  scale_fill_r2dii_sector(sectors = c("automotive", "oil&gas", "power")) +
  theme_2dii() +
  facet_wrap(~sector)
```

* Use `scale_colour_r2dii_tech()` and `scale_fill_r2dii_tech()` to apply the
2DII colour palette for technologies (see `?scale_colour_2dii_tech()` to find 
out what are the available colour labels).

```{r}
technologies <- c("coalcap", "oilcap", "gascap", "hydrocap", "renewablescap")

data <- market_share %>%
  filter(
    metric == "projected",
    sector == "power",
    region == "global",
    year %in% c(2020, 2025)
  ) %>%
  mutate(technology = factor(technology, levels = technologies)) %>%
  arrange(technology)

ggplot(data, aes(x = factor(year), y = production, fill = technology)) +
  geom_col() +
  scale_fill_r2dii_tech("power", technologies) +
  facet_wrap(~technology) +
  theme_2dii()
```

## Funding

This project has received funding from the [European Union LIFE
program](https://wayback.archive-it.org/12090/20210412123959/https://ec.europa.eu/easme/en/) and the [International Climate
Initiative
(IKI)](https://www.international-climate-initiative.com/en/details/project/measuring-paris-agreement-alignment-and-financial-risk-in-financial-markets-18_I_351-2982).
The Federal Ministry for the Environment, Nature Conservation and Nuclear Safety
(BMU) supports this initiative on the basis of a decision adopted by the German
Bundestag. The views expressed are the sole responsibility of the authors and do
not necessarily reflect the views of the funders. The funders are not
responsible for any use that may be made of the information it contains.
