---
title: "r2dii.plot"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of r2dii.plot is to help you plot 2DII data in an informative,
beautiful, and easy way. It is designed to work smoothly with other "r2dii"
packages -- [r2dii.data](https://2degreesinvesting.github.io/r2dii.data/),
[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/), and
[r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/). It also
plays well with the [ggplot2](https://ggplot2.tidyverse.org/) package, which
helps you customize your plots.

```{r setup}
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
library(r2dii.data)
library(r2dii.match)
library(r2dii.plot)
library(r2dii.analysis)
```

Your data should have a structure similar to that of the demo datasets in the
r2dii.data package.

```{r}
loanbook <- loanbook_demo
ald <- ald_demo
scenario <- co2_intensity_scenario_demo
region <- region_isos_demo
```

Your workflow involves functions from the packages r2dii.match and
r2dii.analysis.

```{r}
matched <- loanbook %>%
  match_name(ald) %>%
  prioritize() # Remember to validate matches (see `?prioritize`)
```

r2dii.plot supports three kinds of plots -- emission intensity, techmix, and
trajectory. Each plot can be created using two types of functions: `qplot_*()`
for standardized plots and `plot_*()` for 'bare' plots leaving the customization 
to the user. 

For each kind, you'll need to subset the specific rows you want to
plot (for example with `subset()` or `dplyr:filter()`). For details see the 
documented "Requirements" of the argument `data` of each `plot_*()` or 
`qplot_*()` function (e.g. see `?plot_emission_intensity`). If you forget to 
meet the `data` requirements the error message should guide you.

## 'Quick' plots

Quick plots allow you to create a good looking plot without extensive knowledge 
of other R packages. They are perfect to have a quick look into the data and for 
users who are interested in standard PACTA visualizations. Their common 
characteristics are:

* 5 year time span.
* Formatted data labels.
* Formatted axis labels.
* A title (and a subtitle).

Use `qplot_emission_intensity()` with `sda`-like data.

```{r}
matched %>%
  target_sda(ald, co2_intensity_scenario = scenario) %>%
  subset(sector == "cement") %>%
  qplot_emission_intensity()
```

Use `qplot_trajectory()` with `market_share`-like data.

```{r}
matched %>%
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) %>%
  subset(technology == "renewablescap" & region == "global") %>%
  qplot_trajectory()
```

Use `qplot_techmix()` with `market_share`-like data.

```{r}
matched %>%
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) %>%
  subset(
    sector == "power" &
      region == "global" &
      metric %in% c("projected", "corporate_economy", "target_sds")
  ) %>%
  qplot_techmix()
```

## Plots

Plots created with `plot_*()` functions show the data as they are. To customize
these plots you can use three strategies: 1) Use parameters of `plot_*()`, 2)
use `ggplot2` functions, and 3) Modify the input `data`. This section shows the
three strategies applied to different functions of the `plot_*()` family.

The basic output of `plot_emission_intensity()` looks rather unappealing.

```{r}
p <- matched %>%
  target_sda(ald, co2_intensity_scenario = scenario) %>%
  subset(sector == "cement") %>%
  plot_emission_intensity()
p
```

You can use parameters of `plot_emission_intensity()` to replicate features of
`qplot_emission_intensity()`, and add plot labels with `ggplot2::labs()`.

```{r}
p <- matched %>%
  target_sda(ald, co2_intensity_scenario = scenario) %>%
  subset(sector == "cement") %>%
  plot_emission_intensity(convert_label = to_title, span_5yr = TRUE) +
  labs(
    title = "Emission Intensity Trend for the Cement Sector",
    x = "Year",
    y = "Tons of CO2 per Ton of Production Unit"
  )
p
```

You can polish your plot by modifying the input `data` and output "ggplot"
object. Your options are limitless but these are some typical things you may do:

* Change the time span.
* Add custom labels by modifying the column 'metric' and technology of 
`market_share`-like data.
* Add a title and a subtitle. 
* Change x and y axis labels.
* Customize the colours and legend labels with `ggplot2::scale_colour_manual()`
or `r2dii.plot::scale_*()` functions (see [Styling functions]).

For example, here is how you might customize each of the three kinds of plots:

* `plot_emission_intensity()`.

```{r}
data <- sda %>%
  filter(
    sector == "cement",
    year <= 2030
  )

plot_emission_intensity(data) +
  labs(
    title = "Emission intensity plot for cement",
    x = "Time",
    y = "Tons of CO2 per ton of cement produced"
  ) +
  scale_color_manual(
    values = c("#4a5e54", "#a63d57", "#78c4d6", "#f2e06e"),
    labels = c("Proj.", "Corp. Economy", "Target (demo)", "Adj. Scenario (demo)")
  )
```

* `plot_trajectory()`.

```{r}
matched %>%
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) %>%
  filter(
    technology == "renewablescap",
    region == "global",
    year <= 2030
  ) %>%
  mutate(
    label = case_when(
      metric == "projected" ~ "Your Portfolio",
      metric == "corporate_economy" ~ "Benchmark (Corp. Economy)",
      metric == "target_sds" ~ "Sustainable Development Scen.",
      metric == "target_sps" ~ "Stated Policies Scen.",
      metric == "target_cps" ~ "Current Policy Scen.",
      TRUE ~ metric
    )
  ) %>%
  plot_trajectory() +
  scale_x_continuous(n.breaks = 3) +
  labs(
    title = "Portfolio Scenario Alignment for Renewables Technology",
    x = "Year",
    y = "Production normalized to start year"
  ) +
  theme(plot.margin = unit(c(0.5, 6, 0.5, 1), "cm"))
```

* `plot_techmix()`.

```{r}
data <- market_share %>%
  filter(
    metric %in% c("projected", "corporate_economy", "target_sds"),
    sector == "power",
    region == "global",
    year >= 2021,
    year <= 2040 # custom time range
  ) %>%
  mutate(
    label = case_when(
      metric == "projected" ~ "Your Portfolio",
      metric == "corporate_economy" ~ "Corporate Economy Benchmark",
      metric == "target_sds" ~ "SDS Scenario"
    )
  )

plot_techmix(data) +
  scale_fill_manual(
    values = c("black", "brown", "grey", "blue", "green4"),
    labels = paste(c("Coal", "Oil", "Gas", "Hydro", "Renewables"), "Cap.")
  )
```

## Styling functions

A number of functions allow you to customize your plots using the 2DII style:

* Use `theme_2dii()` to display non-data content.

```{r}
data <- market_share %>%
  filter(
    metric == "projected",
    sector == "power",
    region == "global",
    year %in% c(2020, 2025)
  )

ggplot(data, aes(x = factor(year), y = production)) +
  geom_col() +
  facet_wrap(~technology) +
  theme_2dii()
```

* Use `scale_colour_r2dii()` and `scale_fill_r2dii()` to apply the 2DII colour
palette.

```{r}
data <- market_share %>%
  filter(
    metric != "corporate_economy",
    sector == "power",
    region == "global",
    technology == "renewablescap"
  )

ggplot(data, aes(x = year, y = production, color = metric)) +
  geom_line() +
  scale_colour_r2dii(labels = c("dark_blue", "green", "orange", "ruby_red")) +
  theme_2dii()
```

* Use `scale_colour_r2dii_sector()` and `scale_fill_r2dii_sector()` to apply
the 2DII colour palette for sectors.

```{r}
data <- market_share %>%
  filter(
    metric == "projected",
    region == "global",
    year %in% c(2020, 2025)
  ) %>%
  group_by(sector, year) %>%
  summarise(production = sum(production))

ggplot(data, aes(x = factor(year), y = production, fill = sector)) +
  geom_col() +
  scale_fill_r2dii_sector(sectors = c("automotive", "oil&gas", "power")) +
  theme_2dii() +
  facet_wrap(~sector)
```

* Use `scale_colour_r2dii_fill()` and `scale_fill_r2dii_fill()` to apply the
2DII colour palette for technologies.

```{r}
data <- market_share %>%
  filter(
    metric == "projected",
    sector == "power",
    region == "global",
    year %in% c(2020, 2025)
  ) %>%
  mutate(
    technology = factor(
      technology,
      levels = c("coalcap", "oilcap", "gascap", "hydrocap", "renewablescap")
    )
  ) %>%
  arrange(technology)

technologies <- c("coalcap", "oilcap", "gascap", "hydrocap", "renewablescap")
ggplot(data, aes(x = factor(year), y = production, fill = technology)) +
  geom_col() +
  scale_fill_r2dii_tech("power", technologies) +
  facet_wrap(~technology) +
  theme_2dii()
```
