---
title: "r2dii.plot X versus Y"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of r2dii.plot is to help you plot 2DII data in an informative,
beautiful, and easy way.

## Installation

You can install the development version of r2dii.plot from
[GitHub](https://github.com/2DegreesInvesting/r2dii.plot) with:

```r
# install.packages("devtools")
devtools::install_github("2DegreesInvesting/r2dii.plot")
```

## Example

The r2dii.plot package is designed to work smoothly with other "r2dii" packages
-- [r2dii.data](https://2degreesinvesting.github.io/r2dii.data/),
[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/), and
[r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/). It also
plays well with the popular packages [dplyr](https://www.tidyverse.org/) and
[ggplot2](https://ggplot2.tidyverse.org/), which help you customize your plots.

```{r setup}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(r2dii.plot)
```

Your data typically comes from the output of two functions in the r2dii.analysis
package:
[`target_sda()`](https://2degreesinvesting.github.io/r2dii.analysis/reference/target_sda.html)
and
[`target_market_share()`](https://2degreesinvesting.github.io/r2dii.analysis/reference/target_market_share.html).
Here you'll use two example datasets that come with r2dii.plot.

```{r}
sda

market_share
```

r2dii.plot currently supports three kinds of plots: `plot_timeline*()`,
`plot_techmix*()`, and `plot_trajectory*()`. Each plot has specific requirements
about the main input -- passed to the first argument `data`. To meet those
requirements we currently provide two experimental sets of functions -- or
Application User Interfaces (APIs), "X" and "Y":

* With the "X" API you meet the `data` requirements with `dplyr::filter()`.

* With the "Y" API you meet the `data` requirements with dedicated "preparation"
functions (`prep_*Y()`), for example `prep_timelineY()`:

This small example highlights the essence of the differences between the "X" and
"Y" API.

* "X" API

```{r}
data <- sda
prep <- filter(data, sector == "cement")
plot_timelineX(prep, extrapolate = TRUE)
```

* "Y" API

```{r}
data <- sda
prep <- prep_timelineY(data, sector_filter = "cement", extrapolate = )
plot_timelineY(prep)
```

For full examples see the dedicated articles [r2dii.plot
X](https://2degreesinvesting.github.io/r2dii.plot/articles/articles/r2dii-plot-X.html)
and [r2dii.plot
Y](https://2degreesinvesting.github.io/r2dii.plot/articles/articles/r2dii-plot-Y.html).
