---
title: "r2dii.plot"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of r2dii.plot is to help you plot 2DII data in an informative,
beautiful, and easy way. It is designed to work smoothly with other "r2dii"
packages -- [r2dii.data](https://2degreesinvesting.github.io/r2dii.data/),
[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/), and
[r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/). It also
plays well with the [ggplot2](https://ggplot2.tidyverse.org/) package, which
helps you customize your plots.

```{r setup}
library(ggplot2)
library(r2dii.data)
library(r2dii.match)
library(r2dii.plot)
library(r2dii.analysis)
```

Your data should have a structure similar to that of the demo datasets in the
r2dii.data package.

```{r}
loanbook <- loanbook_demo
ald <- ald_demo
scenario <- co2_intensity_scenario_demo
region <- region_isos_demo
```

Your workflow involves functions from the packages r2dii.match and
r2dii.analysis.

```{r}
matched <- loanbook |>
  match_name(ald) |>
  prioritize()  # Remember to validate matches (see `?prioritize`)
```

## 'Quick' plots

Quick plots allow you to create a good looking plot without extensive knowledge 
of other R packages. They are perfect to have a quick look into the data and for 
users who are interested in standard PACTA outcomes. Their common characteristics
are:

* 5 year time span,
* 'Pretty' labels,
* Standard colour palette
* 'Pretty' axis labels
* Title (and subtitle)

Use `qplot_emission_intensity()` with `sda`-like data.

```{r}
matched |>
  target_sda(ald, co2_intensity_scenario = scenario) |>
  subset(sector == "cement") |>
  qplot_emission_intensity()
```

Use `qplot_trajectory()` with `market_share`-like data.

```{r}
matched |>
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) |>
  subset(technology == "renewablescap" & region == "global") |>
  qplot_trajectory()
```

Use `qplot_techmix()` with `market_share`-like data.

```{r}
matched |>
  target_market_share(ald, scenario = scenario_demo_2020, region_isos = region) |>
  subset(
    sector == "power" & 
      region == "global" &
      metric %in% c("projected", "corporate_economy", "target_sds")
  ) |>
  qplot_techmix()
```

## 
