---
title: "r2dii.plot Y"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of r2dii.plot is to help you plot 2DII data in an informative,
beautiful, and easy way.

## Installation

You can install the development version of r2dii.plot from
[GitHub](https://github.com/2DegreesInvesting/r2dii.plot) with:

```r
# install.packages("devtools")
devtools::install_github("2DegreesInvesting/r2dii.plot")
```

## Example

The r2dii.plot package is designed to work smoothly with other "r2dii" packages
-- [r2dii.data](https://2degreesinvesting.github.io/r2dii.data/),
[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/), and
[r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/). It also
plays well with the popular packages [dplyr](https://www.tidyverse.org/) and
[ggplot2](https://ggplot2.tidyverse.org/), which help you customize your plots.

```{r setup}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(r2dii.plot)
```

Your data typically comes from the output of two functions in the r2dii.analysis
package:
[`target_sda()`](https://2degreesinvesting.github.io/r2dii.analysis/reference/target_sda.html)
and
[`target_market_share()`](https://2degreesinvesting.github.io/r2dii.analysis/reference/target_market_share.html).
Here you'll use two example datasets that come with r2dii.plot.

```{r}
sda

market_share
```

Before plotting, you need to prepare your data. The following sections show the
preparation and plotting workflow for the three kinds of plots that r2dii.plot
supports: timeline, techmix, and trajectory.

### Timeline

The timeline plot uses `sda`-like data. First prepare it with
`prep_timelineY()`, then plot it with `plot_timelineY()`.

Use `prep_timelineY()` to prepare the data. You'll need to subset the data to
meet the requirements of the `plot_timelineY()`, which you can do with the
argument `sector_filter`. If you forget, the error message should guide you.

```{r error = TRUE}
# Fails
prep_timelineY(sda)

# Works
cement <- prep_timelineY(sda, sector_filter = "cement")
# Same
cement <- prep_timelineY(filter(sda, sector == "cement"))

cement
```

* `plot_timelineY()` defaults to recoding `line_name` to title case, and allows
custom recoding via a data frame passed to the argument `specs`.

```{r}
# Default recoding
plot_timelineY(cement)

# Custom recoding
custom <- tribble(
                ~line_name,                  ~label,        ~hex,
               "projected",                 "Proj.",   "#4a5e54",
       "corporate_economy",         "Corp. Economy",   "#a63d57",
             "target_demo",         "Target (demo)",   "#78c4d6",
  "adjusted_scenario_demo",  "Adj. Scenario (demo)",   "#f2e06e",
)

plot_timelineY(cement, specs = custom)
```

### Techmix

TODO

### Trajectory

TODO
