---
title: "r2dii.plot X"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of r2dii.plot is to help you plot 2DII data in an informative,
beautiful, and easy way.

## Installation

You can install the development version of r2dii.plot from
[GitHub](https://github.com/2DegreesInvesting/r2dii.plot) with:

```r
# install.packages("devtools")
devtools::install_github("2DegreesInvesting/r2dii.plot")
```

## Example

The r2dii.plot package is designed to work smoothly with other "r2dii" packages
-- [r2dii.data](https://2degreesinvesting.github.io/r2dii.data/),
[r2dii.match](https://2degreesinvesting.github.io/r2dii.match/), and
[r2dii.analysis](https://2degreesinvesting.github.io/r2dii.analysis/). It also
plays well with the popular packages [dplyr](https://www.tidyverse.org/) and
[ggplot2](https://ggplot2.tidyverse.org/), which help you customize your plots.

```{r setup}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(r2dii.plot)
```

Your data typically comes from the output of two functions in the
r2dii.analysis package:
[`target_sda()`](https://2degreesinvesting.github.io/r2dii.analysis/reference/target_sda.html)
and
[`target_market_share()`](https://2degreesinvesting.github.io/r2dii.analysis/reference/target_market_share.html).
Here you'll use two example datasets that come wiht r2dii.plot.

```{r}
sda

market_share
```

### Timeline

Use `plot_timelineX()` with `sda`-like data. You'll need to pick the
specific rows you want to plot. For details see the documented "Requirements" of
the argument `data`, or try intuitively and let the error messages guide you.

```{r error = TRUE}
plot_timelineX(sda)
```

The error message suggests you must first pick only one value of `sector`, for
example "cement".

```{r}
sda %>% 
  filter(sector == "cement") %>% 
  plot_timelineX()
```

Great! You can now polish you plot. Your options are limitless but these are
some things you may do:

* Pick a narrower time range.
* Recode the legend.
* Extrapolate the timeline.
* Add a title. 

```{r}
to_title <- function(x) gsub("_", " ", tools::toTitleCase(x))

sda %>% 
  filter(sector == "cement", year >= 2020) %>% 
  mutate(emission_factor_metric = to_title(emission_factor_metric)) %>% 
  plot_timelineX(extrapolate = TRUE) + labs(title = "Timeline plot")
```

### Techmix

Use `plot_techmixX()` with `market_share`-like data. Again, learn which rows to
pick by reading the documented "Requirements" of the argument `data`, or by trial and error.

```{r error = TRUE}
unique(market_share$metric)
chosen_metrics <- c("projected", "corporate_economy", "target_sds")

market_share %>%
  filter(
    metric %in% chosen_metrics,
    sector == "power",
    region == "global"
  ) %>% 
  plot_techmixX()
```

### Trajectory

Use `plot_trajectoryX()` with `market_share`-like data. Again, learn which rows
to pick by reading the documented "Requirements" of the argument `data`, or by
trial and error.

```{r}
market_share %>% 
  filter(
    sector == "power", 
    region == "global", 
    technology == "renewablescap"
  ) %>% 
  plot_trajectoryX()
```
